---
output:
  word_document: default
  html_document: default
---
---
title: "Data Analysis Report: One-Way ANOVA (PlantGrowth)"
author: "Bonn Andrei M. Buenaventura"
date: "2026-04-24"
output:
  word_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(ggplot2)
library(car)
library(dplyr)
library(effectsize)

data("PlantGrowth")
```

# Introduction

A one-way Analysis of Variance (ANOVA) is used to determine whether there are statistically significant differences between the means of three or more independent groups.

# Dataset and Problem

The dataset used is *PlantGrowth*, which examines plant weights under different treatment conditions.

* Independent Variable: group (ctrl, trt1, trt2)
* Dependent Variable: weight

The objective is to determine whether the mean plant weight differs across treatment groups.

# Hypotheses

* **H0**: μ₁ = μ₂ = μ₃ (all group means are equal)
* **HA**: At least one group mean is different

# Checking of Assumptions

## Assumption #1: Continuous Dependent Variable

**Remark:** The dependent variable *weight* is continuous.

## Assumption #2: Categorical Independent Variable

**Remark:** The independent variable *group* consists of three independent categories (ctrl, trt1, trt2).

## Assumption #3: Independence of Observations

**Remark:** Observations are assumed independent since each plant is measured separately.

## Assumption #4: No Significant Outliers

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Weight Distribution by Group",
       x = "Group",
       y = "Weight")
```

**Remark:** No significant outliers were detected based on visual inspection of boxplots.

## Assumption #5: Normality

```{r}
normality <- PlantGrowth %>%
  group_by(group) %>%
  summarise(p_value = shapiro.test(weight)$p.value)

normality
```

**Remark:** Data is approximately normally distributed for each group since all Shapiro-Wilk test p-values are greater than 0.05.

## Assumption #6: Homogeneity of Variance

```{r}
leveneTest(weight ~ group, data = PlantGrowth)
```

**Remark:** Homogeneity of variances is satisfied if p > 0.05.

# Descriptive Statistics

```{r}
desc <- PlantGrowth %>%
  group_by(group) %>%
  summarise(
    N = n(),
    Mean = mean(weight),
    SD = sd(weight)
  )

desc
```

# Computation

## One-Way ANOVA

```{r}
res.aov <- aov(weight ~ group, data = PlantGrowth)
anova_result <- summary(res.aov)

F_value <- anova_result[[1]][["F value"]][1]
p_value <- anova_result[[1]][["Pr(>F)"]][1]

anova_result
```

## Effect Size

```{r}
eta <- eta_squared(res.aov)
eta_value <- eta$Eta2[1]

eta
```

## Post Hoc Test

```{r}
tukey <- TukeyHSD(res.aov)
tukey
```

# APA Reporting

A one-way ANOVA was conducted to determine whether plant weight differed across treatment groups (ctrl, trt1, trt2).

There were no outliers, as assessed by boxplot inspection; data were normally distributed for each group (Shapiro-Wilk test, p > .05); and homogeneity of variances was satisfied (Levene’s test, p > .05).

There was a statistically significant difference in mean plant weight between groups,
F(2, 27) = `r round(F_value, 3)`,
p = `r round(p_value, 4)`,
η² = `r round(eta_value, 3)`.

Post hoc comparisons using Tukey’s HSD test indicated which groups differed significantly.

Therefore, the null hypothesis was rejected.
