---
title: "Two-Way ANOVA Analysis: Fertilizer Blend and Crop Yield"
author: "Bonn Andrei M. Buenaventura"
date: "2026-04-29"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(car)
library(knitr)
```

# Introduction

This study evaluates the effectiveness of three fertilizer blends (**X, Y, and Z**) across four crop types (**Wheat, Corn, Soy, and Rice**).

The purpose of this analysis is to determine whether fertilizer blend, crop type, and their interaction significantly affect crop yield.

**Null Hypothesis ($H_0$):**  
There is no significant interaction effect between fertilizer blend and crop type on yield.

---

# Data Entry

```{r}
data <- data.frame(
  Blend = rep(c("X", "Y", "Z"), each = 20),
  Crop = rep(rep(c("Wheat", "Corn", "Soy", "Rice"), each = 5), times = 3),
  Yield = c(
    123,156,112,100,168,
    128,150,174,116,109,
    166,178,187,153,195,
    151,125,117,155,158,

    135,130,176,120,155,
    175,132,120,187,184,
    140,145,159,131,126,
    167,183,142,167,168,

    156,180,147,146,193,
    186,138,178,176,190,
    185,206,188,165,188,
    175,173,154,191,169
  )
)

data$Blend <- as.factor(data$Blend)
data$Crop <- as.factor(data$Crop)
```

---

# Descriptive Statistics

The table below shows the mean and standard deviation of crop yield for each fertilizer-crop combination.

```{r}
data %>%
  group_by(Blend, Crop) %>%
  summarise(
    Mean = round(mean(Yield), 2),
    SD = round(sd(Yield), 2),
    .groups = "drop"
  ) %>%
  kable(caption = "Descriptive Statistics by Blend and Crop")
```

---

# Assumption Testing

## Normality of Residuals

```{r}
model <- aov(Yield ~ Blend * Crop, data = data)

shapiro.test(residuals(model))
```

## Homogeneity of Variances

```{r}
leveneTest(Yield ~ Blend * Crop, data = data)
```

---

# Two-Way ANOVA Results

```{r}
summary(model)
```

---

# Interaction Plot

```{r fig.width=7, fig.height=5}
interaction.plot(
  x.factor = data$Crop,
  trace.factor = data$Blend,
  response = data$Yield,
  fun = mean,
  type = "b",
  col = c("red", "blue", "darkgreen"),
  pch = 19,
  lwd = 2,
  xlab = "Crop Type",
  ylab = "Mean Yield",
  trace.label = "Fertilizer Blend"
)
```

---

# APA Style Report

## Results

A 3 × 4 two-way ANOVA was conducted to examine the effects of fertilizer blend and crop type on crop yield.

Results showed:

- a significant main effect of **fertilizer blend**
- a significant main effect of **crop type**
- a borderline significant interaction effect between fertilizer blend and crop type

Because the interaction p-value is approximately **.050**, the null hypothesis was rejected at the .05 significance level.

This suggests that fertilizer effectiveness depends on the crop being treated.

---

# Conclusion

The analysis indicates that both fertilizer blend and crop type influence agricultural yield.

Among the three formulations, **Blend Z** showed the strongest and most consistent performance, especially for **Soy** and **Corn**.

---

# GitHub Repository
https://github.com/bonnandrei24/Statistical-Theory-Buenaventura-Repository/blob/main/FA12_BUENAVENTURA.Rmd 
