---
title: "Two-Way ANOVA Analysis: Fertilizer Blend and Crop Yield"
author: "Bonn Andrei M. Buenaventura"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(car)
library(effectsize)
library(ggpubr)
library(rstatix)
```

# Introduction

This study examines whether crop yield differs depending on:

1. **Fertilizer Blend** (Blend X, Blend Y, Blend Z)
2. **Crop Type** (Wheat, Corn, Soy, Rice)

A two-way ANOVA was conducted to determine:
- Main effect of fertilizer blend
- Main effect of crop type
- Interaction effect between fertilizer blend and crop type

**Null Hypothesis:** There is no significant interaction effect on yield between fertilizer blend and crop type.

**GitHub Repository:** [Insert GitHub Link Here]

# Data Entry

```{r}
data <- data.frame(
  Blend = rep(c("X", "Y", "Z"), each = 20),
  Crop = rep(rep(c("Wheat", "Corn", "Soy", "Rice"), each = 5), times = 3),
  Yield = c(
    123,156,112,100,168,
    128,150,174,116,109,
    166,178,187,153,195,
    151,125,117,155,158,

    135,130,176,120,155,
    175,132,120,187,184,
    140,145,159,131,126,
    167,183,142,167,168,

    156,180,147,146,193,
    186,138,178,176,190,
    185,206,188,165,188,
    175,173,154,191,169
  )
)

data$Blend <- as.factor(data$Blend)
data$Crop <- as.factor(data$Crop)
head(data)
```

# Descriptive Statistics

```{r}
desc_stats <- data %>%
  group_by(Blend, Crop) %>%
  summarise(
    Mean = mean(Yield),
    SD = sd(Yield),
    n = n()
  )

desc_stats
```

# Visualization

## Boxplot

```{r}
ggplot(data, aes(x = Crop, y = Yield, fill = Blend)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Crop Yield by Fertilizer Blend and Crop",
    x = "Crop Type",
    y = "Yield"
  )
```

## Interaction Plot

```{r}
interaction.plot(
  x.factor = data$Crop,
  trace.factor = data$Blend,
  response = data$Yield,
  fun = mean,
  type = "b",
  legend = TRUE,
  ylab = "Mean Yield",
  xlab = "Crop",
  trace.label = "Blend"
)
```

## Mean Plot

```{r}
ggline(data,
       x = "Crop",
       y = "Yield",
       color = "Blend",
       add = c("mean_se"),
       palette = "jco") +
  theme_minimal()
```

# Assumption Checking

## Normality of Residuals

```{r}
model <- aov(Yield ~ Blend * Crop, data = data)
residuals_model <- residuals(model)

shapiro.test(residuals_model)
qqnorm(residuals_model)
qqline(residuals_model)
```

## Homogeneity of Variances

```{r}
leveneTest(Yield ~ Blend * Crop, data = data)
```

# Two-Way ANOVA

```{r}
anova_results <- Anova(lm(Yield ~ Blend * Crop, data = data), type = 3)
anova_results
```

## Effect Size

```{r}
eta_squared(model, partial = TRUE)
```

# Post Hoc Tests

## Tukey HSD for Blend

```{r}
TukeyHSD(model, "Blend")
```

## Tukey HSD for Crop

```{r}
TukeyHSD(model, "Crop")
```

# APA Style Report

## Results

A two-way ANOVA was conducted to examine the effects of fertilizer blend (Blend X, Blend Y, Blend Z) and crop type (Wheat, Corn, Soy, Rice) on crop yield.

There was a main effect of fertilizer blend, *F*(2, 48) = [insert F], *p* = [insert p], partial η² = [insert eta].

There was a main effect of crop type, *F*(3, 48) = [insert F], *p* = [insert p], partial η² = [insert eta].

There was an interaction effect between fertilizer blend and crop type, *F*(6, 48) = [insert F], *p* = [insert p], partial η² = [insert eta].

If the interaction is significant, the effect of fertilizer blend on crop yield depends on crop type.

## Interpretation

- Reject the null hypothesis if interaction p < .05.
- Fail to reject the null hypothesis if interaction p >= .05.

# Conclusion

This analysis evaluates whether fertilizer effectiveness depends on crop type.

**GitHub Repository:** https://github.com/USERNAME/REPOSITORY-NAME
