---
title: "Two-Way Mixed ANOVA Analysis: Cholesterol Over Time"
author: "Bonn Andrei M. Buenaventura"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(rstatix)
library(ggpubr)
library(ez)
library(car)
library(emmeans)
library(knitr)
library(kableExtra)
```

# Introduction

This study examined whether cholesterol levels differed between two
brands of margarine across three time points:

- Before treatment
- After 4 weeks
- After 8 weeks

A two-way mixed ANOVA was conducted.

- **Between-subject factor:** Margarine Brand
- **Within-subject factor:** Time

## Null Hypothesis

There is no significant difference in cholesterol levels between the two
brands of margarine over the three time points.

---

# Dataset

```{r}
chol_data <- data.frame(
  ID = 1:18,
  Before = c(6.42, 6.76, 6.56, 4.80, 8.43, 7.49, 8.05, 5.05, 5.77,
             3.91, 6.77, 6.44, 6.17, 7.67, 7.34, 6.85, 5.13, 5.73),
  After4weeks = c(5.83, 6.20, 5.83, 4.27, 7.71, 7.12, 7.25, 4.63, 5.31,
                  3.70, 6.15, 5.59, 5.56, 7.11, 6.84, 6.40, 4.52, 5.13),
  After8weeks = c(5.75, 6.13, 5.71, 4.15, 7.67, 7.05, 7.10, 4.67, 5.33,
                  3.66, 5.96, 5.64, 5.51, 6.96, 6.82, 6.29, 4.45, 5.17),
  Margarine = c("B","B","B","A","B","A","B","A","B",
                "A","B","B","A","A","A","B","A","B")
)

head(chol_data)
```

---

# Data Preparation

```{r}
chol_data$Margarine <- as.factor(chol_data$Margarine)

data_long <- chol_data %>%
  pivot_longer(
    cols = c(Before, After4weeks, After8weeks),
    names_to = "Time",
    values_to = "Cholesterol"
  )

data_long$Time <- factor(
  data_long$Time,
  levels = c("Before", "After4weeks", "After8weeks")
)

data_long$ID <- as.factor(data_long$ID)

head(data_long)
```

---

# Descriptive Statistics

```{r}
desc_stats <- data_long %>%
  group_by(Margarine, Time) %>%
  summarise(
    Mean = mean(Cholesterol),
    SD = sd(Cholesterol),
    N = n(),
    .groups = "drop"
  )

kable(desc_stats, digits = 2)
```

---

# Visualization

## Boxplot

```{r}
ggplot(data_long,
       aes(x = Time, y = Cholesterol, fill = Margarine)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Cholesterol Levels by Margarine Brand Across Time",
    x = "Time",
    y = "Cholesterol"
  )
```

## Interaction Plot

```{r}
ggplot(data_long,
       aes(x = Time,
           y = Cholesterol,
           color = Margarine,
           group = Margarine)) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  theme_minimal() +
  labs(
    title = "Interaction Plot of Margarine Brand and Time",
    x = "Time",
    y = "Mean Cholesterol"
  )
```

---

# Assumption Checking

## Normality Test

```{r}
normality <- data_long %>%
  group_by(Margarine, Time) %>%
  shapiro_test(Cholesterol)

kable(normality, digits = 4)
```

## Homogeneity of Variance

```{r}
leveneTest(Cholesterol ~ Margarine * Time, data = data_long)
```

---

# Two-Way Mixed ANOVA

```{r}
anova_results <- ezANOVA(
  data = data_long,
  dv = Cholesterol,
  wid = ID,
  within = Time,
  between = Margarine,
  type = 3,
  detailed = TRUE
)

anova_results
```

---

# Post Hoc Tests

## Time Comparisons

```{r}
pairwise_time <- data_long %>%
  pairwise_t_test(
    Cholesterol ~ Time,
    paired = TRUE,
    p.adjust.method = "bonferroni"
  )

kable(pairwise_time, digits = 4)
```

## Margarine Comparisons

```{r}
pairwise_brand <- data_long %>%
  pairwise_t_test(
    Cholesterol ~ Margarine,
    paired = FALSE,
    p.adjust.method = "bonferroni"
  )

kable(pairwise_brand, digits = 4)
```

---

# Estimated Marginal Means

```{r}
model_lm <- lm(Cholesterol ~ Margarine * Time, data = data_long)
emmeans(model_lm, ~ Margarine * Time)
```

---

# APA Report

## Results

A two-way mixed ANOVA was conducted to determine the effects of margarine
brand and time on cholesterol levels.

The analysis revealed:

- Main effect of Margarine: see ANOVA output
- Main effect of Time: see ANOVA output
- Margarine × Time interaction: see ANOVA output

### APA Format

Replace with actual values:

> A two-way mixed ANOVA revealed a significant/non-significant main
effect of time, *F*(df1, df2) = XX.XX, *p* = .XXX.

> The main effect of margarine brand was significant/non-significant,
*F*(df1, df2) = XX.XX, *p* = .XXX.

> The interaction between margarine brand and time was
significant/non-significant, *F*(df1, df2) = XX.XX, *p* = .XXX.

---

# Conclusion

The null hypothesis was rejected/not rejected depending on the obtained
p-values.

#GitHub Repository Link
