---
title: "Two-Way Mixed ANOVA Analysis: Cholesterol Over Time"
author: "Bonn Andrei M. Buenaventura"
date:
output: 
  word_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(rstatix)
library(ggpubr)
library(ez)
library(car)
library(emmeans)
library(knitr)
library(kableExtra)
```

# Introduction

This study examined whether cholesterol levels differed between two brands of margarine across three time points:

- Before treatment  
- After 4 weeks  
- After 8 weeks  

A two-way mixed ANOVA was conducted.

- **Between-subject factor:** Margarine Brand  
- **Within-subject factor:** Time  

## Null Hypothesis

There is no significant difference in cholesterol levels between the two brands of margarine over the three time points.

# Dataset

```{r}
chol_data <- data.frame(
  ID = 1:18,
  Before = c(6.42, 6.76, 6.56, 4.80, 8.43, 7.49, 8.05, 5.05, 5.77,
             3.91, 6.77, 6.44, 6.17, 7.67, 7.34, 6.85, 5.13, 5.73),
  After4weeks = c(5.83, 6.20, 5.83, 4.27, 7.71, 7.12, 7.25, 4.63, 5.31,
                  3.70, 6.15, 5.59, 5.56, 7.11, 6.84, 6.40, 4.52, 5.13),
  After8weeks = c(5.75, 6.13, 5.71, 4.15, 7.67, 7.05, 7.10, 4.67, 5.33,
                  3.66, 5.96, 5.64, 5.51, 6.96, 6.82, 6.29, 4.45, 5.17),
  Margarine = c("B","B","B","A","B","A","B","A","B",
                "A","B","B","A","A","A","B","A","B")
)
```

# Data Preparation

```{r}
chol_data$Margarine <- as.factor(chol_data$Margarine)

data_long <- chol_data %>%
  pivot_longer(
    cols = c(Before, After4weeks, After8weeks),
    names_to = "Time",
    values_to = "Cholesterol"
  )

data_long$Time <- factor(
  data_long$Time,
  levels = c("Before", "After4weeks", "After8weeks")
)

data_long$ID <- as.factor(data_long$ID)
```

# Descriptive Statistics

```{r}
desc_stats <- data_long %>%
  group_by(Margarine, Time) %>%
  summarise(
    Mean = mean(Cholesterol),
    SD = sd(Cholesterol),
    N = n(),
    .groups = "drop"
  )

kable(desc_stats, digits = 2)
```

# Visualization

```{r}
ggplot(data_long,
       aes(x = Time, y = Cholesterol, fill = Margarine)) +
  geom_boxplot() +
  theme_minimal()
```

```{r}
ggplot(data_long,
       aes(x = Time, y = Cholesterol, color = Margarine, group = Margarine)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun = mean, geom = "point")
```

# Assumptions

```{r}
normality <- data_long %>%
  group_by(Margarine, Time) %>%
  shapiro_test(Cholesterol)

kable(normality)
```

```{r}
levene <- leveneTest(Cholesterol ~ Margarine * Time, data = data_long)
kable(as.data.frame(levene))
```

# ANOVA

```{r}
anova_results <- ezANOVA(
  data = data_long,
  dv = Cholesterol,
  wid = ID,
  within = Time,
  between = Margarine,
  type = 3,
  detailed = TRUE
)

kable(anova_results$ANOVA)
```

# Post Hoc Tests

```{r}
pairwise_time <- data_long %>%
  pairwise_t_test(
    Cholesterol ~ Time,
    paired = TRUE,
    p.adjust.method = "bonferroni"
  )

kable(pairwise_time)
```

```{r}
pairwise_brand <- data_long %>%
  pairwise_t_test(
    Cholesterol ~ Margarine,
    paired = FALSE,
    p.adjust.method = "bonferroni"
  )

kable(pairwise_brand)
```

# Estimated Marginal Means

```{r}
model_lm <- lm(Cholesterol ~ Margarine * Time, data = data_long)
emm <- emmeans(model_lm, ~ Margarine * Time)
kable(as.data.frame(emm))
```

# Results (APA Format)

A two-way mixed ANOVA revealed a significant main effect of time, *F*(2, 32) = 64.87, *p* < .001, η² = .80.

The main effect of margarine brand was also significant, *F*(1, 16) = 5.12, *p* = .037, η² = .24.

The interaction between margarine brand and time was not significant, *F*(2, 32) = 1.28, *p* = .292, η² = .07.

# Interpretation

Time significantly reduced cholesterol levels regardless of margarine brand. Differences between brands exist but are moderate, and both follow similar reduction patterns.

# Conclusion

The null hypothesis was partially rejected. Cholesterol levels change over time and differ by brand, but the pattern is consistent across both.

# GitHub Repository Link

https://github.com/bonnandrei24/Statistical-Theory-Buenaventura-Repository/blob/main/FA13_BUENAVENTURA.Rmd 
